---
title: "ProcessEEGData"
output: html_document
date: "2024-07-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

~/Library/CloudStorage/GoogleDrive-brichard21@gmail.com/.shortcut-targets-by-id/1mhzB6Bm-yAwNrG1a7MkMaBWWAMHzl07L/binocularitynoise/existingscripts/Analysis expt 1/R/reProcessEEG.Rmd

```{r}
library(R.matlab)
library(pracma)
library(addAlpha)
library(wesanderson)

colours <- wes_palette("IsleofDogs1")
coloursT <- add.alpha(colours)

is.pc <- .Platform$OS.type == "windows"
```

```{r}
theParticipants <- c("BJ", "BR", "DHB", "FL", "GV", "JM", "KR", "LEH", "LH", "MM", "MVH", "RZ", "SO", "TNL", "XX", "YUN")
```

```{r}
if (is.pc){
  theBlockDirectory <- "G:/.shortcut-targets-by-id/1mhzB6Bm-yAwNrG1a7MkMaBWWAMHzl07L/binocularitynoise/Experiment1data/"
} else {
  theBlockDirectory <- "~/Google Drive/My Drive/YORK/binocularitynoise/Experiment1data/"
}
```

MM (participant 10) has an extra trigger in block 2 at trial 34 (35 with the first 0 trigger) - 35 (trial 35 happened 1 sample after trial 34) - manually fixed by removing trial 34.
XX has bad data and was removed from the bigger analysis file.

```{r}
P <- 1

stimulusConditions <- matrix(NA, nrow= 39, ncol = 4)
theEEGData <- array(NA, dim = c(64,10000,39,4))
theFFTData <- array(NA, dim = c(64,10000,39,4))

blankTime <- 4000
duration <- 10000

for (BLOCK in 1:4){
  theFile <- readMat(paste0(theBlockDirectory, theParticipants[P], "/EEG_", theParticipants[P], "BLOCK", BLOCK, ".mat"))
  
  if (P == 15 | P == 16){
    stimulusConditionsTemp <- as.numeric(unlist(theFile$EEG[,,1]$event["type",1,]))
    stimulusConditionsTemp[1] <- 0
    triggerTimes <- unlist(theFile$EEG[,,1]$event["latency",1,])*1000
  } else {
    stimulusConditionsTemp <- unlist(theFile$EEG[,,1]$triggers["type",1,])
    triggerTimes <- unlist(theFile$EEG[,,1]$triggers["offset",1,])
    if (P == 10 & BLOCK == 2){
      stimulusConditionsTemp <- stimulusConditionsTemp[-35]
      triggerTimes <- triggerTimes[-35]
    }
  }
  
  electrodes <- unlist(theFile$EEG["label",,1]$label)
  
  triggerTimes <- triggerTimes[!(stimulusConditionsTemp %in% c(0, 99))] 
  stimulusConditions[, BLOCK] <- stimulusConditionsTemp[!(stimulusConditionsTemp %in% c(0, 99))]
  for (event in 1:length(stimulusConditions[, BLOCK])){
    for (channel in 1:(length(electrodes)-2)){
      theEEGData[channel,,event,BLOCK] <- theFile$EEG[,,1]$data[channel, (triggerTimes[event]+blankTime):(triggerTimes[event]+duration+blankTime-1)]
      theFFTData[channel,,event,BLOCK] <- fft(theFile$EEG[,,1]$data[channel, (triggerTimes[event]+blankTime):(triggerTimes[event]+duration+blankTime-1)])/10000
    }
  }
}
```

```{r}
theChannels <- c("Oz", "POz", "O1", "O2")
eData <- theEEGData[electrodes[1:64] %in% theChannels,,,]
eFFTData <- theFFTData[electrodes[1:64] %in% theChannels,,,]
```

```{r, eval = FALSE}

block <- 2
plot(1:10000,eData[1,,trial,block]-mean(eData[1,,trial,block]), type = "l")
lines(1:10000, eData[2,,trial,block]-mean(eData[2,,trial,block]), col = "red")
lines(1:10000, eData[3,,trial,block]-mean(eData[3,,trial,block]), col = "blue")
lines(1:10000, eData[4,,trial,block]-mean(eData[4,,trial,block]), col = "magenta")

plot(1:100, abs(eFFTData[1,2:101,trial,block]), type = "l", col = coloursT[4])
abline(v = 30)
lines(1:100, abs(eFFTData[2,2:101,trial,block]), col = coloursT[4])
lines(1:100, abs(eFFTData[3,2:101,trial,block]), col = coloursT[4])
lines(1:100, abs(eFFTData[4,2:101,trial,block]), col = coloursT[4])
lines(1:100, abs(apply(eFFTData[,2:101,trial,block],2,mean)), col = colours[2], lwd = 2)

c(electrodes[1:64] %in% theChannels)*1:64
```

```{r}
eAvgFFTData <- apply(eFFTData,c(2,3,4),mean)
eAvgEEGData <- apply(eData, c(2,3,4), mean)
avgSNRCondition <- array(NA, dim = c(200,13))
for (condition in 1:13){
  theIDX <- which(stimulusConditions == condition, arr.ind = TRUE)
  
  conditionEEGTemp <- matrix(NA,nrow=10000,ncol =12)
  conditionFFTTemp <- matrix(NA,nrow=10000,ncol =12)
  
  for (rep in 1:12){
    conditionEEGTemp[,rep] <- eAvgEEGData[,theIDX[rep,1], theIDX[rep,2]]
    conditionFFTTemp[,rep] <- eAvgFFTData[,theIDX[rep,1], theIDX[rep,2]]
  }
  
  tempxy <- data.frame(Re(conditionFFTTemp[31,]),Im(conditionFFTTemp[31,]))
  D <- sqrt(mahalanobis(tempxy,colMeans(tempxy),cov(tempxy)))
  outIDX <- which(D < 3)
  conditionFFT <- conditionFFTTemp[ ,outIDX]
  conditionEEG <- conditionEEGTemp[ ,outIDX]
  
  SNRData <- matrix(0, nrow = 200, ncol = dim(conditionFFT)[2])
  SNRData2 <- matrix(0, nrow = 200, ncol = dim(conditionFFT)[2])
  
  for (trial in 1:dim(conditionFFT)[2]){
    for (f in 5:200){
      SNRData[f,trial] <- abs(conditionFFT[f, trial])/mean(abs(conditionFFT[c((f-5),(f+5)), trial]))
      
    }
  }
  
  avgSNRCondition[,condition] <- apply(SNRData, 1, mean)
}

# if (is.pc){
#   save(avgSNRCondition, file = paste0("C:/Users/Bruno/Box/Binocularity Noise/Reprocessed/reprocessedData_P", theParticipants[P], ".RData"))
# } else {
#   save(avgSNRCondition, file = paste0("~/Library/CloudStorage/Box-Box/Binocularity Noise/Reprocessed/reprocessedData_P", theParticipants[P], ".RData"))
# }
```

```{r}
posterColours <- c("#333E47", "#6a757c", "#e31936")

condition <- 6
theIDX <- which(stimulusConditions == condition, arr.ind = TRUE)
conditionEEGTemp <- matrix(NA,nrow=10000,ncol =12)
conditionFFTTemp <- matrix(NA,nrow=10000,ncol =12)

for (rep in 1:12){
  conditionEEGTemp[,rep] <- eAvgEEGData[,theIDX[rep,1], theIDX[rep,2]]
  conditionFFTTemp[,rep] <- eAvgFFTData[,theIDX[rep,1], theIDX[rep,2]]
}

tempxy <- data.frame(Re(conditionFFTTemp[31,]),Im(conditionFFTTemp[31,]))
D <- sqrt(mahalanobis(tempxy,colMeans(tempxy),cov(tempxy)))
outIDX <- which(D < 3)
conditionFFT <- conditionFFTTemp[ ,outIDX]
conditionEEG <- conditionEEGTemp[ ,outIDX]

binoc <- apply(abs(conditionFFT), 1, mean)
binocEEG <- apply(conditionEEG, 1, mean)
binocEEG[1:1000] <- binocEEG[1:1000] - mean(binocEEG[1:1000])


# png("~/Library/CloudStorage/Box-Box/Binocularity Noise/Paper_test/Figures/EEG.png", width = 6.5, height = 3.5, res = 300,units = "in", bg = "transparent")
yLimit <- range(pretty(range(pretty(range(binocEEG[1:1000])))))
plot(x = NULL, y = NULL, xlim = c(0,1000), ylim = yLimit, axes = FALSE, ann = FALSE)
axis(1, at = pretty(c(0,1000)), tck = -.015, lwd = 2, labels = FALSE)
axis(2, at = pretty(yLimit), tck = -.015, lwd = 2, labels = FALSE)
mtext(pretty(c(0,1000)),1, at = pretty(c(0,1000)), line = .4)
mtext(pretty(yLimit), 2, at = pretty(yLimit), line = .4, las = 1)
abline(h = 0, lwd = 2, col = "gray75")
lines(1:1000, binocEEG[1:1000], lwd = 2, col = posterColours[1])
mtext("Time (ms)", 1, font = 2, line = 1.5)
mtext(bquote(bold("Electrical Potential ("*mu*"V)")), 2, font = 2, line = 2)
lines(1:1000, Real(tes[1:1000])/1000, col = "red")
# dev.off()

png("~/Library/CloudStorage/Box-Box/Binocularity Noise/Paper_test/Figures/SSVEPAmp.png", width = 8, height = 5, res = 300,units = "in", bg = "transparent")
yLimit <- c(0,1)
plot(x = NULL, y = NULL, xlim = c(0,100), ylim = yLimit, axes = FALSE, ann = FALSE)
axis(1, at = pretty(c(0,100)), tck = -.015, lwd = 2, labels = FALSE)
axis(2, at = pretty(yLimit), tck = -.015, lwd = 2, labels = FALSE)
mtext(pretty(c(0,100))/10,1, at = pretty(c(0,100)), line = .4)
mtext(pretty(yLimit), 2, at = pretty(yLimit), line = .4, las = 1)
abline(v = c(30, 60,90), lwd = 2, col = "gray75")
text(c(31,61,91), 2.5, c("3Hz", "6Hz", "9Hz"), adj = 0)
lines(1:100, binoc[2:101], lwd = 2, col = posterColours[1])
mtext("Frequency (Hz)", 1, font = 2, line = 1.5)
mtext(bquote(bold("SSVEP Amplitude ("*mu*"V)")), 2, font = 2, line = 2)
dev.off()
```

```{r}
binocEEG <- apply(conditionEEG, 1, mean)

theFilter<- rep(0, 10000)
theFilter[c(31,61,91,121,151,181)] <- 1

fData <- fft(binocEEG)
fData[1] <- 0
filtered <- abs(fData) * theFilter
tes <- fft(filtered * exp(1i*angle(fData)), inverse = TRUE)

plot(Real(tes[1:1000])/1000)
```


```{r}
condition <- 5
theIDX <- which(stimulusConditions == condition, arr.ind = TRUE)
conditionEEGTemp <- matrix(NA,nrow=10000,ncol =12)
conditionFFTTemp <- matrix(NA,nrow=10000,ncol =12)

for (rep in 1:12){
  conditionEEGTemp[,rep] <- eAvgEEGData[,theIDX[rep,1], theIDX[rep,2]]
  conditionFFTTemp[,rep] <- eAvgFFTData[,theIDX[rep,1], theIDX[rep,2]]
}

tempxy <- data.frame(Re(conditionFFTTemp[31,]),Im(conditionFFTTemp[31,]))
D <- sqrt(mahalanobis(tempxy,colMeans(tempxy),cov(tempxy)))
outIDX <- which(D < 3)
conditionFFT <- conditionFFTTemp[ ,outIDX]
conditionEEG <- conditionEEGTemp[ ,outIDX]

antiphase <- apply(abs(conditionFFT), 1, mean)
plot(antiphase[2:100])
```

```{r, fig.width=8}
coloursT <- add.alpha(colours)

yLimit <- c(-40,40)
xLimit <- c(0,1000)
# png("~/Library/CloudStorage/Box-Box/Binocularity Noise/Paper/Figures/ElectrodePotential.png", width = 9.5, height = 5.5, res = 300, units = "in", bg = "transparent")
plot(x = NULL, y = NULL, xlim = xLimit, ylim = yLimit, axes = FALSE, ann = FALSE)
axis(1, at = pretty(xLimit), tck = -.015, lwd = 2, labels = FALSE, pos = 0)
axis(1, at = pretty(xLimit), tck = .015, lwd = 2, labels = FALSE, pos = 0)
axis(2, at =  pretty(yLimit), tck = -.015, lwd = 2, labels = FALSE,)
mtext(pretty(yLimit), 2, at = pretty(yLimit), las = 1, line = .5)
text(pretty(xLimit), -7, pretty(xLimit))
text(1000, -40, "Time (ms)", font = 1, adj = 1)
mtext(bquote("Electrical Potential"~" ("~mu~"V)"), 2, line = 2, adj = 0.5)

lines(eData[1,3000:4000,theIDX[2,1], theIDX[2,2]]-mean(eData[1,,theIDX[2,1], theIDX[2,2]]), lwd = 2, col = coloursT[4])
lines(eData[2,3000:4000,theIDX[2,1], theIDX[2,2]]-mean(eData[2,,theIDX[2,1], theIDX[2,2]]), lwd = 2, col = coloursT[3])
lines(eData[3,3000:4000,theIDX[2,1], theIDX[2,2]]-mean(eData[3,,theIDX[2,1], theIDX[2,2]]), lwd = 2, col = coloursT[2])
lines(eData[4,3000:4000,theIDX[2,1], theIDX[2,2]]-mean(eData[4,,theIDX[2,1], theIDX[2,2]]), lwd = 2, col = coloursT[1])
legend("topleft", legend = theChannels, lwd = 2, col = coloursT[4:1], bty = "n", cex = .8)
# dev.off()
```

```{r}
yLimit <- c(0,3)
xLimit <- c(0,100)
# png("~/Library/CloudStorage/Box-Box/Binocularity Noise/Paper/Figures/FFTSNR.png", width = 9, height = 3, res = 300, units = "in", bg = "transparent")

par(mfrow = c(1,2), mai = c(.2,.6,0,.25), omi = c(.4,.1,.05,.05))
plot(x = NULL, y = NULL, xlim = xLimit, ylim = yLimit, axes = FALSE, ann = FALSE)
axis(1, at = pretty(xLimit)+1, tck = -.015, lwd = 2, labels = FALSE)
axis(2, at =  pretty(yLimit), tck = -.015, lwd = 2, labels = FALSE)
mtext(pretty(yLimit),2,at = pretty(yLimit), line = .4, las = 1)
mtext(pretty(xLimit)/10,1,at = pretty(xLimit)+1, line = .4, las = 1)
mtext("Frequency (Hz)", 1, font = 1, line = 1.5)
mtext(bquote("Amplitude"~" ("~mu~"V)"), 2, line = 2)
rect(31-5,0,31+5,3, col = coloursT[5], border = NA)
rect(61-5,0,61+5,3, col = coloursT[5], border = NA)
rect(91-5,0,91+5,3, col = coloursT[5], border = NA)
abline(v = c(31, 61, 91), lty = 2, col = colours[5])

lines(2:100, abs(eAvgFFTData[2:100,theIDX[2,1], theIDX[2,2]]),col = colours[6], lwd = 2)
text(c(32,62,92), 3.05, paste0(c(3,6,9), "Hz"), adj = 0, cex = .8)


yLimit <- c(1,4)
xLimit <- c(0,100)
plot(x = NULL, y = NULL, xlim = xLimit, ylim = yLimit, axes = FALSE, ann = FALSE)
axis(1, at = pretty(xLimit)+1, tck = -.015, lwd = 2, labels = FALSE)
axis(2, at =  pretty(yLimit), tck = -.015, lwd = 2, labels = FALSE)
mtext(pretty(yLimit),2,at = pretty(yLimit), line = .4, las = 1)
mtext(pretty(xLimit)/10,1,at = pretty(xLimit)+1, line = .4, las = 1)
mtext("Frequency (Hz)", 1, font = 1, line = 1.5)
mtext("SNR", 2, line = 2.5)
abline(v = c(31, 61, 91), lty = 2, col = colours[5])
text(c(32,62,92), 4.05, paste0(c(3,6,9), "Hz"), adj = 0, cex = .8)

lines(2:100, avgSNRCondition[2:100,6],col = colours[6], lwd = 2)
# dev.off()
```

```{r}

```




```{r, fig.width=15}
xLim = c(0,200)
theCond <- matrix(c(1,2,3,13,NA, 4:12, NA), nrow = 3, ncol = 5, byrow = TRUE)
par(mfrow = c(3,5), mai = c(.2,.2,.2,.1),omi = c(.25,.25,.25,.25))
yLim <- c(1,max(range(pretty(range(avgSNRCondition)))))
for (II in 1:3){
  for (C in theCond[II,]){
    plot(x = NULL, y = NULL, xlim = xLim, ylim = yLim, axes = FALSE, ann = FALSE)
    if (!is.na(C)){
      axis(2, at = pretty(yLim), lwd = 2, tck = -.015, labels = FALSE)
      
      if (II == 3){
        mtext(pretty(xLim)/10, 1, at = pretty(xLim), line = .3)
        mtext("Frequency (Hz)", 1, font = 2, line = 1.5)
      }
      if (C == 1 | C == 4 | C == 9){
        mtext(pretty(yLim), 2, at = pretty(yLim), line = .3, las = 1)
        mtext("SNR", 2, font = 2, line = 2)
      }
      
      abline(v = 31, lty = 2, col = colours[3], lwd = 2)
      abline(v = 61, lty = 2, col = colours[3], lwd = 2)
      abline(v = 91, lty = 2, col = colours[3], lwd = 2)
      abline(v = 121, lty = 2, col = colours[3], lwd = 2)
      abline(v = 151, lty = 2, col = colours[3], lwd = 2)
      abline(v = 181, lty = 2, col = colours[3], lwd = 2)
      
      text(31.5, 6, "3Hz", adj = 0)
      text(61.5, 6, "6Hz", adj = 0)
      text(91.5, 6, "9Hz", adj = 0)
      text(121.5,6, "12Hz", adj = 0)
      text(151.5,6, "15Hz", adj = 0)
      text(181.5,6, "18Hz", adj = 0)
      
      lines(1:200, avgSNRCondition[,C], lwd = 4, col = colours[1])
      mtext(paste0("Condition ", C), 3, line = 0, adj = 0)
      axis(1, at = pretty(xLim), lwd = 2, tck = -.015, labels = FALSE)
    }
  }
}
```











